\documentclass{article}

\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage[noend]{algpseudocode}


\begin{document}
    \title{Tutorial 1}
    \date{18-07-2019}
    \author{Deepank Agrawal (17CS30011)}
  \maketitle
  
  \section{Problem Statement}
    $A[1..m]$ and $B[1..n]$ are two 1D arrays containing $m$ and $n$ integers 
    respectively, where $m\le n$.
    We need to construct a sub-array $C[1..m]$ of $B$ such that the
    \newline\textbf{expression} $\sum\limits_{i=1}^{m} \big|A[i]-C[i]\big|$ is minimized.
  
  \section{Recurrences}
    
    To solve the problem, we can design a Dynamic Programming(DP) algorithm. The
    formulation of the DP is: \\
    Suppose that for any $i \, \epsilon \,[1, m]$ and $k \, \epsilon
    \,[1, n]$, $ i \le k$, we have already scanned arrays $A[1..i-1]$ and
    constructed $C[1..i-1]$, which is the optimal sub-array. Now, for $A[i]$, on scanning 
    any $B[k]\,\forall \, k\,\epsilon\,[1, n]$, the following two cases arise for 
    construction of the optimal sub-array C:
    \begin{enumerate}
        \item $B[k]$ is included in C.
        \item $B[k]$ is not included in C.
    \end{enumerate}
    If $B[k]$ is included in C, then $C[1..i-1]\,\cup\,B[k]$ should be the so-far optimal
    sub-array. Else, $B[k]$ is not included and $C[1..i-1]$ is the optimal sub-array.
    This is decided by whether or not including $B[k]$ minimizes $M[i][k]$ i.e., whether
    $\big|A[i]-B[k]\big|\,+\,M[i-1][k-1]\,\textless\,M[i][k-1]$ or not.
    
    Let's consider an array $M[1..m][1..n]$, where $M[i][k]$ stores the minimal value of 
    the expression for arrays $A[1..i]$ and $B[1..k]$.
    Now the recurrence relation for the DP, $\forall i\, \epsilon \,[0,
    m],\,k\,\epsilon\,[0, n]$, can be defined as:
    
    \begin{equation}
        M[i][k]=
        \begin{cases}
            0, & \text{$i=0$}\\
            \infty, & \text{$i\,\textless\,k$}\\
            min\left\{\big|A[i]-B[k]\big|\,+\,M[i-1][k-1],\,M[i][k-1]\right\}, & 
            \text{$k\geq i$}\\
        \end{cases}
    \end{equation}

    So, $M[m][n]$ will have the final minimum value of the expression.
    
  \section{Algorithm}
    
    Now that Optimal Substructure has been defined, let's design the algorithm. Before 
    that, we can see there are \textbf{overlapping sub-problems} (e.g. $M[3][5]$ will
    be required for both $M[4][6]$ and $M[3][6]$). So, \textbf{memoization} will be used.
    
    To solve the problem, we store the minimal value of the expression $\forall 
    i, \,k,\, i\le k$ using equation (1) and along-side store $B[k]$ to be included in 
    the optimal sub-array. For this we construct a matrix $M[0..m][0..n]$, where 
    $M[i][k]$ stores the minimal value of the expression for arrays $A[1..i]$ and 
    $B[1..k]$ and an array $C[0..m]$ to store the included $B[k]$ values. Bottom-up
    approach will be used to fill matrix $M$.

    The optimal sub-array $C$ is updated when $M[i][k]\,\ne\,M[i][k-1]$ in a top-down
    fashion. For this, iterate matrix $M\,:\,(m,\,n)\,\rightarrow\,(1,\,1)$ 
    and check for $M[i][k]\,\ne\,M[i][k-1]$ condition.\\
    
    \newline\newline \textbf{Pseudocode:}
    
    \begin{enumerate}
        \item \textbf{initialize $M[1..m][0..n]$ with $\infty$ and $M[0][0..n]$ with $0$}
        \item \textbf{for $i : 1\,\rightarrow\,m\,$ do} 
        \quad\quad\Comment{\textit{Construct M}}
        \item \quad\textbf{for $k : i\,\rightarrow\,n\,$ do}
        \item \quad\quad\textbf{$M[i][k]\,\leftarrow\,min\left\{\big|A[i]-B[k]\big| + 
        M[i-1][k-1], M[i][k-1]\right\}$}
        \item \quad\textbf{end for}
        \item \textbf{end for}
        \item \textbf{$i\,\leftarrow\,m\,,\,k\,\leftarrow\,n$}
        \item \textbf{while $i\,\ne\,0$ do} 
        \quad\quad\Comment{\textit{Construct C}}
        \item \quad\textbf{if $M[i][k]\,\ne\,M[i][k-1]$, then do}
        \item \quad\quad\textbf{$C[i]\,\leftarrow\,B[k]$}
        \item \quad\quad\textbf{$i\,\leftarrow\,i-1$}
        \item \quad\textbf{end if}
        \item \quad\textbf{$k\,\leftarrow\,k-1$}
        \item \textbf{end while}
    \end{enumerate}
    \pagebreak
  \section{Demonstration}
  
    \begin{enumerate}
        \item Let $A\,=\,[9,\,10]\,\&\,B\,=\,[4,\,9,\,14]$\\
        Now, $M[1][1]\,=\,\text{min}\{\big|A[1]-B[1]\big|+M[0][0],\,M[1][0]\}$\\
        or, $M[1][1]\,=\,\text{min}\{\big|9-4\big|+0,\,\infty\}$
        or, $M[1][1]\,=\,\text{min}\{5,\,\infty\}\,=\,5$\\
        and, $C[1]\,=\,B[1]\,=\,4$.\\
        Similarly, the matrix $M[0..m][0..n]$ will be:\\
        \newline$M$ = \begin{bmatrix}
                0 & 0 & 0 & 0 \\
                \infty & 5 & 0 & 0 \\
                \infty & \infty & 6 & 4 
              \end{bmatrix}\\
        \newline Here, since $M[2][3] \ne M[2][2]$ so, $C[2]\,=\,B[3]\,=\,14$\\
        and, $M[1][2] \ne M[1][1]$ so, $C[1]\,=\,B[2]\,=\,9$\\
        Hence, $C\,=\,[9,\,14]$ is the optimal sub-array.
        \newline\newline\item Let $A\,=\,[2,\,7,\,2]\,\&\,B\,=\,[5,\,3,\,6,\,8]$\\
        As above, constructing the matrix $M[0..m][0..n]$:\\
        \newline$M$ = \begin{bmatrix}
                0 & 0 & 0 & 0 & 0 \\
                \infty & 3 & 1 & 1 & 1 \\
                \infty & \infty & 7 & 2 & 2 \\
                \infty & \infty & \infty & 11 & 8
              \end{bmatrix}\\
        \newline Here, $M[3][4]\,\ne\,M[3][3],\,M[2][3]\,\ne\,M[2][2],\,M[1][2]\,\ne\,M[1]
        [1]$\\
        So, $C[3]\,=\,B[4]\,=8,\,C[2]=\,B[3]\,=6,\,C[1]\,=\,B[2]\,=3$\\
        Hence, $C\,=\,[3,\,6,\,8]$ is the optimal sub-array.
        \newline\newline\item Let 
        $A\,=\,[9,\,10,\,12]\,\&\,B\,=\,[7,\,6,\,9,\,8,\,9,\,12]$\\
        As above, constructing the matrix $M[0..m][0..n]$:\\
        \newline$M$ = \begin{bmatrix}
                0 & 0 & 0 & 0 & 0 & 0 & 0 \\
                \infty & 2 & 2 & 0 & 0 & 0 & 0 \\
                \infty & \infty & 6 & 3 & 2 & 1 & 1 \\
                \infty & \infty & \infty & 9 & 7 & 5 & 2 \\
              \end{bmatrix}\\
        \newline Here, $M[3][6]\,\ne\,M[3][5],\,M[2][5]\,\ne\,M[2][4],\,M[1][3]\,\ne\,M[1]
        [2]$\\
        So, $C[3]\,=\,B[6]\,=12,\,C[2]=\,B[5]\,=9,\,C[1]\,=\,B[3]\,=9$\\
        Hence, $C\,=\,[9,\,9,\,12]$ is the optimal sub-array.
    \end{enumerate}
    \pagebreak
  \section{Time and space complexities}
  
    \subsection{Time Complexity}
    Let's refer to the above mentioned pseudocode for calculation of time complexity.
    As we can see, from line no. $2\,\&\,3$, the for loops are of $O(m)\,and\,O(n)$
    complexity respectively. Inside the nested for loops, the operation is of 
    $O(1)$ time. Also, from line no. $8$, the while loop is of $O(n)$ time. So, the 
    total time complexity of the algorithm is $O(mn)$ because of the nested loops.
    
    \subsection{Space Complexity}
    As a matrix $M[0..m][0..n]$ is constructed, the space complexity of the algorithm 
    will also be $O(mn)$.\\
    
\end{document}
